This section records the formalism choices, simulator capabilities, and
conventions that apply to every experiment in this repository.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Foundational References}

The tick-counter experiment reproduced in this repository originates
from~\cite{VDW14}.  The CDBoost simulator architecture used to run all
experiments is described in~\cite{VNWD15}.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{CDBoost PDEVS}

CDBoost implements Parallel DEVS (PDEVS) as originally formulated,
following the sequential call/return architecture of~\cite{VNWD15}.
That paper defines how all simulation steps execute in a single thread
with a call stack whose depth is linear in the model hierarchy --- an
important property for resource-constrained deployments.

The original PDEVS formalism does not specify a port system; CDBoost
follows this faithfully.  Coupling is expressed as a list of
\texttt{(sender, receiver)} model pointer pairs; all submodels in a
coupled model share a single message type \texttt{MSG}.  This is a
deliberate design choice, not a restriction.

Multiple simultaneous messages are delivered as a
\textbf{message bag} (type \texttt{std::vector<MSG>}).

When a model is both imminent and receives external input, the
\textbf{confluence function} $\delta_{\text{con}}$ resolves the two transitions.
CDBoost's default (and the convention used in these experiments) is
internal-first:
\[
  \delta_{\text{con}}(s, e, x) =
    \delta_{\text{ext}}\!\left(\delta_{\text{int}}(s),\; 0,\; x\right)
\]

\subsubsection*{Coupled Model Structure}

\texttt{coupled<TIME>} takes four constructor arguments:
\begin{itemize}
  \item \texttt{models} --- all submodel pointers
  \item \texttt{eic}    --- external input coupling (models that receive external input)
  \item \texttt{ic}     --- internal coupling as \texttt{(sender, receiver)} pairs
  \item \texttt{eoc}    --- external output coupling (models whose output leaves the coupled model)
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Message Type}

CDBoost uses \texttt{MSG = std::any} as the single shared message type
across a coupled model.  The only constraint the interface imposes is
that two models connected via IC, EOC, or EIC must agree on the
concrete C++ type stored inside \texttt{std::any} at that connection.
The simulator itself enforces nothing beyond this: all messages travel
as type-erased \texttt{std::any} values.

The concrete type carried inside \texttt{std::any} is an
experiment-level decision.  Choosing it for a particular experiment or
coupling those models with different concrete types is the
responsibility of the modeller, not the simulator.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{TIME Type}

CDBoost is parameterised over a \texttt{TIME} template argument.
Infinity is provided via \texttt{time\_inf<TIME>::value()}; a
specialisation is required for types that do not expose
\texttt{std::numeric\_limits<TIME>::infinity()}.

Common choices:

{\small
\begin{center}
\begin{tabular}{@{}llp{5cm}@{}}
  \toprule
  Type & Representation & Risk \\
  \midrule
  \texttt{double} & IEEE~754 double &
    $1/10$ not exact; errors may break the causality
    chain and are unbounded \\[4pt]
  \texttt{boost::rational<int>} & exact fraction &
    no rounding error; heavier compute \\
  \bottomrule
\end{tabular}
\end{center}}

CDBoost ships \texttt{cdboost/rational\_time.hpp}, which provides a
\texttt{time\_inf} specialisation for \texttt{boost::rational<int>},
so that \texttt{boost::rational} satisfies the TIME requirements
without further wrapping.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Log Format}

CDBoost emits structured \textbf{NDJSON} (Newline-Delimited JSON) to
stdout --- one JSON object per line.  Every record carries a
wall-clock timestamp (\texttt{ts}), a severity level (\texttt{level}),
a structured event name (\texttt{event}), and a human-readable message
(\texttt{msg}).  Events that occur during the simulation additionally
include \texttt{sim\_time}, the logical clock value at that instant.

In these experiments the log will contain:
\begin{itemize}
  \item a \texttt{simulation\_start} record at $t = 0$,
  \item one \texttt{tick} record per output emitted by the counter
        model, carrying the count in \texttt{msg},
  \item a \texttt{simulation\_end} record at the configured time limit,
  \item a \texttt{performance} record reporting total wall time.
\end{itemize}

For the complete event schema, optional fields, and \texttt{jq}
recipes for post-processing, see
\texttt{simulators/cdboost/docs/log-format.md}.
