% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Overview}

The tick-counter model simulates a periodic tick source that is counted
over time and periodically registered and reset.  Two generators
produce events at different rates: a fast generator fires every
$1/10$~s (the \emph{tick}) and a slow generator fires every $1$~s (the
\emph{reset}).  A counter model accumulates incoming ticks and, upon
receiving a reset signal, outputs the current count and resets to zero.
Ideally the counter outputs $10$ at every reset.  The experiment
reveals whether the chosen TIME type preserves this invariant under
floating-point arithmetic.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Model Diagram}

\begin{center}
\begin{tikzpicture}[
  box/.style={draw, rectangle, minimum width=1.8cm, minimum height=0.9cm, align=center},
  >=Stealth
]
  \node[box] (g10) at (0, 0.9) {$G_{1/10}$};
  \node[box] (g1)  at (0,-0.9) {$G_1$};
  \node[box] (k)   at (4.5, 0) {$K$};

  \draw[->] (g10) -- node[above, font=\small] {\texttt{int}: 1} (k);
  \draw[->] (g1)  -- node[below, font=\small] {\texttt{int}: 0} (k);
  \draw[->] (k)   -- node[above, font=\small] {\texttt{int}: count}
                     (6.5, 0) node[right, font=\small] {\texttt{out}};
\end{tikzpicture}
\end{center}

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Coupled Model}

\[
  C = \langle X_C,\, Y_C,\, D,\, \{M_d\},\, \text{EIC},\, \text{EOC},\, \text{IC} \rangle
\]
\[
  D = \{G_{1/10},\; G_{1},\; K\},\quad X_C = \emptyset,\quad Y_C \subseteq \mathbb{N}
\]
\[
  \text{EIC} = \emptyset
  \qquad
  \text{IC} = \{(G_{1/10} \to K),\; (G_{1} \to K)\}
  \qquad
  \text{EOC} = \{(K \to C.\texttt{out})\}
\]

Expected invariant: every output of $K$ equals $10$.

% ─────────────────────────────────────────────────────────────────────────────
\subsubsection*{Atomic Model Definitions}

\paragraph{Generator $G_{1/10}$.}

\[
  G_{1/10} = \langle X,\; Y,\; S,\;
               \delta_{\text{ext}},\; \delta_{\text{int}},\; \delta_{\text{con}},\;
               \lambda,\; ta \rangle
\]
\begin{align*}
  X &= \emptyset \\
  Y &= \mathbb{Z} \\
  S &= \{\bullet\} \\
  \delta_{\text{ext}} &= \bot
    \quad\text{(undefined; } X = \emptyset\text{)} \\
  \delta_{\text{int}}(\bullet) &= \bullet \\
  \delta_{\text{con}} &= \bot
    \quad\text{(undefined; } X = \emptyset\text{)} \\
  \lambda(\bullet) &= 1 \\
  ta(\bullet) &= \tfrac{1}{10}
\end{align*}

\noindent\textit{Implementation.}
\texttt{basic\_models::generator<TIME,\,std::any>}
from namespace \texttt{cdboost::pdevs}; see \texttt{tick\_gen.hpp}.
$S$ is the trivial single state internal to \texttt{generator}.
$Y$ is realised as \texttt{std::any} wrapping \texttt{int};
the concrete value transported is $1$ (tick signal).
No EIC connection --- $X = \emptyset$.

\paragraph{Generator $G_1$.}

\[
  G_1 = \langle X,\; Y,\; S,\;
          \delta_{\text{ext}},\; \delta_{\text{int}},\; \delta_{\text{con}},\;
          \lambda,\; ta \rangle
\]
\begin{align*}
  X &= \emptyset \\
  Y &= \mathbb{Z} \\
  S &= \{\bullet\} \\
  \delta_{\text{ext}} &= \bot \\
  \delta_{\text{int}}(\bullet) &= \bullet \\
  \delta_{\text{con}} &= \bot \\
  \lambda(\bullet) &= 0 \\
  ta(\bullet) &= 1
\end{align*}

\noindent\textit{Implementation.}
\texttt{basic\_models::generator<TIME,\,std::any>}
from namespace \texttt{cdboost::pdevs}; see \texttt{reset\_gen.hpp}.
$Y$ is realised as \texttt{std::any} wrapping \texttt{int};
the concrete value transported is $0$ (reset signal).
No EIC connection --- $X = \emptyset$.

\paragraph{Counter $K$.}

\[
  K = \langle X,\; Y,\; S,\;
       \delta_{\text{ext}},\; \delta_{\text{int}},\; \delta_{\text{con}},\;
       \lambda,\; ta \rangle
\]
\begin{align*}
  X &= \mathbb{Z} \\
  Y &= \mathbb{N} \\
  S &= \mathbb{N} \times \{\text{idle},\; \text{ready}\}
\end{align*}
\begin{align*}
  ta(n,\text{idle})  &= \infty & ta(n,\text{ready}) &= 0 \\[4pt]
  \delta_{\text{ext}}(s,e,x) &=
    \begin{cases}
      (n + |x|,\; \text{idle})  & \forall m \in x: m \neq 0 \\
      (n,\;       \text{ready}) & \exists m \in x: m = 0
    \end{cases} \\[4pt]
  \delta_{\text{int}}(n,\text{ready}) &= (0,\; \text{idle}) \\[4pt]
  \delta_{\text{con}}(s, e, x) &=
    \delta_{\text{ext}}\!\left(\delta_{\text{int}}(s),\; 0,\; x\right) \\[4pt]
  \lambda(n,\text{ready}) &= n
\end{align*}

\noindent\textit{Implementation.}
\texttt{basic\_models::infinite\_counter<TIME,\,std::any>}
from namespace \texttt{cdboost::pdevs}; see \texttt{k\_counter.hpp}.
$S$ is stored as \texttt{(n:~int,~mode)} where \texttt{mode} is an
internal enum \texttt{\{idle,~ready\}}.
$X$ is realised as \texttt{std::any} wrapping \texttt{int}: the concrete
values transported are~$1$ (tick from $G_{1/10}$) and~$0$ (reset from
$G_1$).
$Y$ is realised as \texttt{std::any} wrapping \texttt{int}: the concrete
value transported is the accumulated count $n$.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Simulation Setup}

Both TIME variants are driven by the CDBoost \texttt{runner} with a
fixed target time limit passed to \texttt{runner::runUntil}.  The
runner calls \texttt{collectOutputs} before each
\texttt{advanceSimulation} step and stops when logical time reaches
the limit.  If the two runs produce logs of different length (e.g.\
due to causality breaks in the float run), the longer log is truncated
to the shorter run for comparison.

\paragraph{TIME types exercised.} \texttt{double} and
\texttt{boost::rational<int>}.

% ─────────────────────────────────────────────────────────────────────────────
\subsection*{Expected Observations}

\textbf{Double time:} IEEE~754 cannot represent $1/10$ exactly.
Floating-point error shifts occasional $G_{1/10}$ ticks across a
1-second boundary; $K$ outputs $9$ or $11$.  Errors may break the
causality chain and are unbounded.

\textbf{Rational time:} $G_{1/10}$ and $G_1$ are always simultaneously
imminent at integer seconds; confluence correctly resolves the tie and
$K$ always outputs $10$.
